---
title: "Final Project"
author: "Wayne Maumbe, Meron Gedrago"
date: "December 22, 2017"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
      smooth_scroll: false
    df_print: kable
---

## Load all packages

```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5, echo=FALSE}
library(dplyr)
library(cluster)
library(class)
library(dplyr)
library(cluster)
library(knitr)
library(rpart)
library(caret)
library(MLmetrics)
library(MASS)
library(knncat)
library(readr)
library(tictoc)
library(tidyverse)
library(data.table)
library(gridExtra)
library(rpart.plot)

```
 



## Load data and perform data cleaning

Please delete all "notes" before submission.

Note: CSV/data files should be read assuming they are in the `data` folder. In
other words, load data via `read_csv("data/CSV_NAME.csv")` and not via
`read_csv("/Users/aykim/Documents/MATH495/Final_Project/data/CSV_NAME.csv")`

#Load data and data cleaningfor kNN
```{r, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5, echo=FALSE}
expediatrain <- read_csv("~/Final_Project/data/expediatrain.csv")
test_2 <- read_csv("~/Final_Project/data/test-2.csv")
#we are filtering the train by booking since the test dataset only contains information of booked hotel searches 
Trainset <- filter(expediatrain , is_booking == 1 )
```

#Load data and data cleaning for CART
```{r load_data, message=FALSE, warning=FALSE, echo=FALSE}
train<-read_csv("~/Final_Project/data/expediatrain.csv")
train<-filter(train, train$is_booking==1)
```

```{r set_vartypes, message=FALSE, warning=FALSE, echo=FALSE}
#set variables to reflect appropriate variable type as described in Kaggle
train$site_name<-as.factor(train$site_name)
train$posa_continent<-as.factor(train$posa_continent)
train$user_location_country<-as.factor(train$user_location_country)
train$user_location_region<-as.factor(train$user_location_region)
train$user_location_city<-as.factor(train$user_location_city)
train$is_mobile<-as.factor(train$is_booking)
train$is_package<-as.factor(train$is_package)
train$channel<-as.factor(train$channel)
train$srch_destination_id<-as.factor(train$srch_destination_id)
train$srch_destination_type_id<-as.factor(train$srch_destination_type_id)
train$is_booking<-as.factor(train$is_booking)
train$hotel_continent<-as.factor(train$hotel_continent)
train$hotel_country<-as.factor(train$hotel_country)
train$hotel_market<-as.factor(train$hotel_market)
train$hotel_cluster<-as.factor(train$hotel_cluster)
train$srch_ci<-as.Date(train$srch_ci)
train$srch_co<-as.Date(train$srch_co)
```

```{r shrinkage, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5, echo=FALSE}
#condense the 100 levels of hotel clusters to 9 popularity levels  
Summaryclusters<-tally(group_by(train,hotel_cluster))
Shrink<-mutate(Summaryclusters, popularity=c(1:dim(Summaryclusters)[[1]]))
#define limits for each popularity level
for( i in 1:dim(Shrink)[[1]]){
if (Shrink$n[i]<=20 && Shrink$n>0) {
  Shrink$popularity[i]<-"1"
}
if (Shrink$n[i]<= 40&& Shrink$n[i]>20) {
  Shrink$popularity[i]<-"2"
}
  if (Shrink$n[i]<=60 && Shrink$n[i]>40) {
  Shrink$popularity[i]<-"3"
  }
  if (Shrink$n[i]<= 80 && Shrink$n[i]>60) {
  Shrink$popularity[i]<-"4"
  }
  if (Shrink$n[i]<= 100 && Shrink$n[i]>80) {
  Shrink$popularity[i]<-"5"
  }
  if (Shrink$n[i]<= 115&& Shrink$n[i]>100) {
  Shrink$popularity[i]<-"6"
  }
  if (Shrink$n[i]<= 140&& Shrink$n[i]>115) {
  Shrink$popularity[i]<-"7"
  }
  if (Shrink$n[i]<= 160&& Shrink$n[i]>140) {
  Shrink$popularity[i]<-"8"
  }
  if (Shrink$n[i]>160) {
  Shrink$popularity[i]<-"9"
  }
  
}
#assign to every point of hotel cluster the corresponding popularity level
train<-mutate(train, popularity=c(1:dim(train)[[1]]))
for (j in 1:dim(Shrink)[[1]]) {
 for (i in which(train$hotel_cluster==Shrink$hotel_cluster[j])) {
   train$popularity[i]<-Shrink$popularity[j]
  } 
}
 #download and rename dataset for ease of plotting and data handling for shiny app
write_csv(train,"trainpop.csv")
trainpop<-train
```

```{r date_calc, message=FALSE, warning=FALSE, echo=FALSE}
#calculate and append length of stay 
train$popularity<-as.factor(train$popularity)
train<-mutate( train, date_diff=c(1:dim(train)[[1]]))
train$date_diff <- train$srch_co- train$srch_ci
train$date_diff<-as.numeric(train$date_diff)
```

```{r date_cleaning, message=FALSE, warning=FALSE, echo=FALSE}
#create seperate Date dataframe with only complete cases to allow for usage in anova methods
Dates<-as.data.frame(cbind(train$date_diff, train$popularity))
Summarydate<-tally(group_by(Dates,train$date_diff))
Dates<-mutate(Dates, is_package=trainpop$is_package)
Dates<-mutate(Dates, is_booking=trainpop$is_booking)
Dates<-mutate(Dates, hotel_continent=trainpop$hotel_continent)
Dates<-mutate(Dates, hotel_location=trainpop$hotel_country)
Dates<-mutate(Dates, posa_continent=trainpop$posa_continent)
indx<-c(which(train$date_diff<0))
Dates<-Dates[-indx,]
Dates<-Dates[complete.cases(Dates), ]
Summarydate<-tally(group_by(Dates,Dates$V1))
Dates$V2<-as.factor(Dates$V2)
Dates<-mutate(Dates, logdiff=log(Dates$V1))
levels(Dates$posa_continent)[1]<-"Africa"
levels(Dates$posa_continent)[2]<-"Europe"
levels(Dates$posa_continent)[3]<-"N.America"
levels(Dates$posa_continent)[4]<-"S.America"
levels(Dates$posa_continent)[5]<-"Asia"
levels(Dates$is_package)[1]<-"No"
levels(Dates$is_package)[2]<-"Yes"
Dates[-which(Dates$logdiff==Inf)] = NA
Datesnaless<-Dates[-which(Dates$logdiff==-Inf),]
Datesnaless<-Datesnaless[complete.cases(Datesnaless),]
```



## EDA visualizations and tables

Note: If you had to illustrate using no modelling but only graphs and tables which
variables have the most predictive power, which would you include?

* Perform a cross-validation on only the final/ultimate model used for your
submission.
* The "score" in question should be the same as used to compute the Kaggle
leaderboard. In other words, your estimated score should be roughly equal to the
score returned by Kaggle after your submission.

```{r hue_plus_outcome_plot, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5, echo=FALSE}
#plot of the outcome variable as-is with color showing shrinkage scheme
cols <- c("9" = "#5C0E0E", "6" = "red2", "5" = "red", "2" = "lightpink","4" = "#FF1F1F","3" = "#FC4545", "7" = "#CE0B0B","8" = "#8B0000","1" = "#F6CACA")
ggplot(Shrink, aes(x=hotel_cluster, y=n, color=popularity))+geom_point(position = "dodge")+ xlab("hotel cluster") +ylab("number of searches")+scale_color_manual("hotel popularity levels", values = cols)+theme(axis.text.x = element_text(angle = 90, size = 4))+ theme(legend.title.align=9 ,panel.background = element_rect(fill = NA),panel.grid.major = element_line(colour = "grey90"))
```

```{r anova, message=FALSE, warning=FALSE}
#anova analysis of predictor variables 
aovdiff<-aov(lm(logdiff~is_package+V2+posa_continent+is_package:V2+is_package:posa_continent+is_package:V2:posa_continent, data = Datesnaless,na.action=na.exclude))
summary(aovdiff)
```

## Crossvalidation of ultimate model

Note: Hardcode your crossvalidation here i.e. do not use built-in crossvalidation
options.

#Crossvalidation for kNN 

```{r, warning=FALSE, echo=FALSE}
 
# Consider a range of k values.  
k_vector <- c(1,3,5,7,9,11,13,15)
 
# Save results here
results <- data_frame(
  k_value = k_vector,
  MAPK = 0
)
for(i in 1:length(k_vector)){
  for(j in 1:5){
    # Set up training & validation (AKA "pretend" test) set based on folds
    train_set <- train_subset %>% 
      sample_frac(.2*j)
    validation_set <- anti_join(train_subset,train_set)
  
    
    # Fit model
    model_formula12 <- as.formula(hotel_cluster ~ is_package + hotel_continent +          hotel_market +  hotel_country )
    model  <-  knn3(model_formula12, data=train_set, k = k_vector[i], use.all = TRUE)
 

    # Get predictions
      predictions <- predict(model, newdata =validation_set)  %>%
      as_tibble() %>% 
      pull(y)
    
    # Compute MSE and save
    results$MAPK[i] <- validation_set %>% 
      mutate(yhat = predictions) %>% 
      summarise(MAPK = mapk(5,validation_set$hotel_cluster,validation_set$yhat)) %>% 
      pull(MAPK)
  }
}
```

#Crossvalidation for CART 
```{r formula, message=FALSE, warning=FALSE, echo=FALSE}
popmodelfom<-as.formula(V2~is_package+posa_continent+logdiff)
```

```{r cv, message=FALSE, warning=FALSE}
#cross validation scheme
traincv <- Datesnaless%>% 
  sample_frac(1) %>% 
  mutate(fold = rep(1:10, length=n()))
# Consider a range of complexity parameters. this was chosen by trial and error
counter<-1:10
cp_vector <-c(0,0.000001*counter)
counter<-1:11
# Save results here
results <- data_frame(
  cp = cp_vector,
  score = 25235542 #place holder
)
for(i in counter){
  for(j in 1:10){
    # Set up training & validation (AKA "pretend" test) set based on folds
   
    train_set <- traincv %>% 
      filter(fold == j)
    validation_set <- traincv %>% 
      filter(fold !=j)
    
    # Fit model
    model <- rpart(popmodelfom, data =train_set,method="class", control = rpart.control(cp=cp_vector[[i]]))
    # Get predictions
    y_hat_model <- model %>% 
    predict(type="class",newdata=validation_set)
    
    # Score/error
    
    # results$score[i]<-Metrics::mapk(5, validation_set$popularity, y_hat) 
    #for the cv analysis we employ accuracy to obtain score instead of required mapk
    results$score[i]<-MLmetrics::Accuracy(y_true = validation_set$V2, y_pred = y_hat_model)
    }
}
optimal <- results %>% 
  arrange(score) %>% 
  slice(1)
optimal %>% 
  kable(digits=3)
cp_star <- optimal$cp
score_star <- optimal$score
ggplot(results, aes(x=cp, y=score)) +
  geom_point() +
  labs(x="complexity parameter", y="score", title="Crossvalidation estimates of Accuracy for CART model") +  geom_vline(xintercept = cp_star, col="red")
```


 

## Create submission

Note: Output a CSV using `write_csv(DATAFRAME_NAME, path="data/SUBMISSION_NAME.csv")`
that is Kaggle submitable. This submission should return a Kaggle score that is
close to your crossvalidated score.

#submission of kNN
```{r, warning=FALSE, echo=FALSE}

k <- 5
#subsetting the Trainset because there were too many ties when predicting using knn
train_subset <- Trainset %>% 
    slice(1:10000)

model_formula12 <- as.formula(hotel_cluster ~ is_package + hotel_continent + hotel_market +  hotel_country )
model_knn12 <-  knn3(model_formula12, data=train_subset, k = k, use.all = TRUE)
#find the probabilities also 
 
predictedvalues <- predict(model_knn12, newdata =  test_2)
 

 predictedclusters <- colnames(predictedvalues)[max.col(predictedvalues,ties.method="first")] 

predicted_clusters <- as.data.frame(predictedclusters)
 
submission1 <-  cbind(test_2$id ,predicted_clusters)
colnames(submission1) <- c("id","hotel_cluster")
write_csv(submission1, "submission.csv")
```

#Final model for CART 

```{r cart plotting, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5, echo=FALSE}
tree<-rpart(popmodelfom, data = Dates, method="class",control = rpart.control(cp =0))
bestcp <- tree$cptable[which.min(tree$cptable[,"xerror"]),"CP"]
tree.cv<-rpart(popmodelfom, data = Datesnaless, method="class",control = rpart.control(cp = bestcp, maxdepth = 5))
tree.pruned <- prune(tree.cv, cp = bestcp)
#rpart.plot::prp(tree.cv, varlen=0 ,faclen=0 , cex = 0.8, extra = 100, type = 1,uniform=T, branch=0, yesno=T)
#plot of tree showing pruned version of model 
rpart.plot::prp(tree.pruned,  varlen=0 ,faclen = 0, type=1, cex = 0.8,extra=100,uniform=T, branch=0, yesno=T)
```

```{r load_test_data,message=FALSE,warning=FALSE, echo=FALSE}
testexpedia<-read.csv("~/Advanced Data Analysis/Rstudio and Git projects/Final_Project/data/test.csv")

```


```{r test_data_cleaning,message=FALSE,warning=FALSE, echo=FALSE}
testexpedia$site_name<-as.factor(testexpedia$site_name)
testexpedia$posa_continent<-as.factor(testexpedia$posa_continent)
testexpedia$user_location_country<-as.factor(testexpedia$user_location_country)
testexpedia$user_location_region<-as.factor(testexpedia$user_location_region)
testexpedia$user_location_city<-as.factor(testexpedia$user_location_city)
testexpedia$is_mobile<-as.factor(testexpedia$is_mobile)
testexpedia$is_package<-as.factor(testexpedia$is_package)
testexpedia$channel<-as.factor(testexpedia$channel)
testexpedia$srch_destination_id<-as.factor(testexpedia$srch_destination_id)
testexpedia$srch_destination_type_id<-as.factor(testexpedia$srch_destination_type_id)
testexpedia$hotel_continent<-as.factor(testexpedia$hotel_continent)
testexpedia$hotel_country<-as.factor(testexpedia$hotel_country)
testexpedia$hotel_market<-as.factor(testexpedia$hotel_market)
testexpedia$srch_ci<-as.Date(testexpedia$srch_ci)
testexpedia$srch_co<-as.Date(testexpedia$srch_co)
testexpedia<-mutate( testexpedia, date_diff=c(1:dim(testexpedia)[[1]]))
testexpedia$date_diff <- testexpedia$srch_co- testexpedia$srch_ci
testexpedia$date_diff<-as.numeric(testexpedia$date_diff)
Dates<-as.data.frame(cbind(testexpedia$date_diff, testexpedia$popularity))
Summarydate<-tally(group_by(Dates,testexpedia$date_diff))
Dates<-mutate(Dates, is_package=testexpedia$is_package)
Dates<-mutate(Dates, hotel_continent=testexpedia$hotel_continent)
Dates<-mutate(Dates, posa_continent=testexpedia$posa_continent)
Summarydate<-tally(group_by(Dates,Dates$V1))
Dates<-mutate(Dates, logdiff=log(Dates$V1))
levels(Dates$is_package)[1]<-"No"
levels(Dates$is_package)[2]<-"Yes"
```
```{r test fitting, message=FALSE, warning=FALSE, echo=TRUE}
y_hat <- tree.cv %>% 
  predict(newdata=Dates, type="class")
```
```{r expand_scheme, message=FALSE,warning=FALSE, echo=FALSE}
propodata<-cbind.data.frame(trainpop$hotel_cluster, trainpop$popularity)
pop1<-filter(propodata,trainpop$popularity==1)
pop1weight<-tally(group_by(pop1,`trainpop$hotel_cluster`))%>%
  as_tibble() 
pop1weight$n<-pop1weight$n/sum(pop1weight$n)
pop2<-filter(propodata,trainpop$popularity==2)
pop2weight<-tally(group_by(pop2,`trainpop$hotel_cluster`))%>%
  as_tibble() 
pop2weight$n<-pop2weight$n/sum(pop2weight$n)
pop3<-filter(propodata,trainpop$popularity==3)
pop3weight<-tally(group_by(pop3,`trainpop$hotel_cluster`))%>%
  as_tibble() 
pop3weight$n<-pop3weight$n/sum(pop3weight$n)
pop4<-filter(propodata,trainpop$popularity==4)
pop4weight<-tally(group_by(pop4,`trainpop$hotel_cluster`))%>%
  as_tibble() 
pop4weight$n<-pop4weight$n/sum(pop4weight$n)
pop5<-filter(propodata,trainpop$popularity==5)
pop5weight<-tally(group_by(pop5,`trainpop$hotel_cluster`))%>%
  as_tibble() 
pop5weight$n<-pop5weight$n/sum(pop5weight$n)
pop6<-filter(propodata,trainpop$popularity==6)
pop6weight<-tally(group_by(pop6,`trainpop$hotel_cluster`))%>%
  as_tibble() 
pop6weight$n<-pop6weight$n/sum(pop6weight$n)
pop7<-filter(propodata,trainpop$popularity==7)
pop7weight<-tally(group_by(pop7,`trainpop$hotel_cluster`))%>%
  as_tibble() 
pop7weight$n<-pop7weight$n/sum(pop7weight$n)
pop8<-filter(propodata,trainpop$popularity==8)
pop8weight<-tally(group_by(pop8,`trainpop$hotel_cluster`))%>%
  as_tibble() 
pop8weight$n<-pop8weight$n/sum(pop8weight$n)
pop9<-filter(propodata,trainpop$popularity==9)
pop9weight<-tally(group_by(pop9,`trainpop$hotel_cluster`))%>%
  as_tibble() 
pop9weight$n<-pop9weight$n/sum(pop9weight$n)
levels(y_hat)
tempy_hat<-as.numeric(y_hat)
```
```{r expand, warning=FALSE, echo=FALSE}
set.seed(1509)
lvlyhat<-unique(tempy_hat)
yhat<-tempy_hat
for (i in 1:length(tempy_hat)) {
yhat[[i]]<-replace(yhat[[i]],yhat[[i]]==lvlyhat[[2]],as.numeric(sample(pop8weight$`trainpop$hotel_cluster`, size=1, prob =pop8weight$n )))
}
for (i in 1:length(tempy_hat)) {
yhat[[i]]<-replace(yhat[[i]],yhat[[i]]==lvlyhat[[1]],as.numeric(sample(pop9weight$`trainpop$hotel_cluster`, size=1, prob =pop9weight$n )))
}
unique(yhat)
tempyhat<-as.factor(yhat)
levels(tempyhat)

```

#submission for CART
```{r submission_filing, message=FALSE, warning=FALSE, echo=FALSE}
submission<-cbind.data.frame(sample_submission$id, tempyhat)
colnames(submission)[1]<-"id"
colnames(submission)[2]<-"hotel_cluster"
write_csv(submission, "submission1.csv")
```



## Citations and references

Note: All citations and references must be included here.



## Supplementary materials

Note: Anything else you've tried that you'd like to include, but isn't essential to
the above, like other EDA's, other modeling approaches you've tried, etc. Please
set the R code chunk `eval=FALSE` here so that default is that R Markdown
doesn't run the code, but a user can flip this switch if they are curious.

```{r, eval=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4.5, echo=FALSE}


```



 
 



